Team Ken FJ! analysis

### 2020-12-24 jeopardy_proper/fj_teamken_analysis.Rmd
move to new folder

---
title: "Team Ken FJ! analysis"
author: "Monica Thieu"
date: "11/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(googledrive)
require(magrittr)
require(tidyverse)
```

@@

@@

>>>>

---
title: "Team Ken FJ! analysis"
author: "Monica Thieu"
date: "11/22/2018"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(googledrive)
require(magrittr)
require(tidyverse)
require(rvest)
# Apparently we need this shit now
options(mc.cores = parallel::detectCores())
```

Jeopardy!, the ultimate knowledge-based game show (in my opinion), lends itself very well to analysis about the game itself.

For the 2019 Jeopardy! All-Star Games, I started doing some rough analyses about what kinds of patterns appear in Jeopardy! games that we might be able to harness for our own use.

What this blog post **is:**

What this blog post **is not:**

@@

I realized that a big piece of information we could recover from the Final Jeopardy! clues that we all answered was _clue categories._

@@

## extra lock 2 game shit

I was going to do a rough grid search across possible score sets to see the conformations that end in a game 2 lock, but the dataframe gets huge very quickly and stuff was running a little too slow for my taste.

```{r, eval = FALSE}
scores <- crossing(team1_game1 = 1:30,
                   team2_game1 = 1:30,
                   team3_game1 = 1:30) %>%
  filter(team1_game1 >= team2_game1, team1_game1 >= team3_game1)
```

## Daily Double extravaganza (as heard on Planet Money)

```{r big ass dd scrapey bitch, eval = FALSE}
seasons_raw <- 25:35 %>%
  paste0("http://www.j-archive.com/showseason.php?season=", .) %>%
  map(~read_html(.)) %>%
  map(~html_nodes(., "a"))

jscrape_raw <- seasons_raw %>%
  tibble(season = 25:35,
         raw = .) %>%
  mutate(text = map(raw, ~html_text(., trim = T)),
         link = map(raw, ~html_attr(., "href"))) %>%
  select(-raw) %>%
  unnest() %>%
  filter(grepl("showgame", link)) %>%
  mutate(web_id = as.integer(str_sub(link, start = -4L)),
         game_id = as.integer(str_sub(text, start = 2L, end = 5L)),
         these_htmls = map(web_id, ~read_html(paste0("http://j-archive.com/showgame.php?game_id=", .x))),
         these_nodes = map(these_htmls, ~html_nodes(.x, ".clue")))

jscrape <- jscrape_raw$these_nodes %>%
  map(~.x %>%
        html_text(trim = T) %>%
        tibble(raw = .) %>%
        # bye bye FJ, for now at least
        slice(-nrow(.))) %>%
  tibble(season = jscrape_raw$season,
         web_id = jscrape_raw$web_id,
         stuff = .) %>%
  filter(map_int(stuff, nrow) == 60) %>%
  unnest() %>%
  separate(raw, into = c("value", "order", "blank1", "blank2", "clue"), sep = "\n", fill = "right") %>%
  select(-starts_with("blank")) %>%
  mutate_all(str_trim, side = "both") %>%
  mutate(web_id = as.integer(web_id)) %>%
  # Splicing clue point value for the empty cells
  group_by(web_id) %>%
  mutate(idx = 1:n()) %>%
  filter(web_id != 3719) %>%
  mutate(column = rep(1:6, times = 10),
         row = rep(1:5, each = 6, times = 2),
         round = rep(c("sj", "dj"), each = 30),
         is_dd = if_else(startsWith(value, "DD"), TRUE, FALSE),
         nspaces_clue = str_count(clue, " ")) %>%
  select(-value, -idx)

write_csv(jscrape, "dd_scrape_25_35.csv")
```

```{r}
jscrape <- read_csv("dd_scrape_25_35.csv")
```

```{r}
jscrape %>%
  group_by(season, web_id, round, row, column) %>%
  summarize(n = sum(is_dd),
            mean_nspaces = mean(nspaces_clue, na.rm = TRUE)) %>%
  mutate(category_length = if_else(mean_nspaces > 10,
                                      "long",
                                      "short",
                                      missing = "short")) %>%
  group_by(round, row, column, category_length) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  mutate(round = recode(round,
                        sj = "Single Jeopardy",
                        dj = "Double Jeopardy"),
         round = fct_rev(round),
         n = na_if(n, 0),
         n = if_else(round == "Double Jeopardy",
                     n/2,
                     as.double(n))) %>%
  ggplot(aes(x = column, y = row, fill = n)) +
  geom_raster() +
  scale_y_reverse() +
  viridis::scale_fill_viridis() +
  # Remove category_length if not conditioning by this
  facet_grid(category_length ~ round) +
  labs(x = "Clue column",
       y = "Clue row",
       title = "Heatmap of DD incidence, most of seasons 25-35",
       subtitle = "DDs don't come up in wordplay categories",
       fill = "# of DDs") +
  theme_bw()
```

```{r}
jscrape %>%
  group_by(web_id, round, column) %>%
  summarize(mean_nspaces = mean(nspaces_clue, na.rm = TRUE),
            contains_dd = sum(is_dd)) %>%
  ungroup() %>%
  mutate(round = recode(round,
                        sj = "Single Jeopardy",
                        dj = "Double Jeopardy"),
         round = fct_rev(round)) %>%
  ggplot(aes(x = mean_nspaces, fill = factor(contains_dd))) +
  geom_density(alpha = 0.5) +
  labs(x = "# spaces per clue (proxy for word count)",
       title = "Category wordiness by DD incidence") +
  facet_grid(~round) +
  theme_bw()
```

```{r}
jscrape %>%
  filter(round == "dj", is_dd) %>%
  select(web_id, row) %>%
  group_by(web_id) %>%
  mutate(index = paste0("row_", 1:n())) %>%
  spread(index, row) %>%
  select(-row_3) %>%
  mutate(row_diff = row_2 - row_1) %>%
  ggplot(aes(x = row_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "ROW distance ($ amount) between DJ DDs") +
  theme_bw()
```

```{r}
jscrape %>%
  filter(round == "dj", is_dd) %>%
  select(web_id, column) %>%
  group_by(web_id) %>%
  mutate(index = paste0("col_", 1:n())) %>%
  spread(index, column) %>%
  # This one game has 3 DDs listed but it must be wrong lol
  select(-col_3) %>%
  mutate(col_diff = abs(col_2 - col_1)) %>%
  ggplot(aes(x = col_diff)) +
  geom_histogram(binwidth = 1) +
  labs(x = "COL distance (categories) between DJ DDs") +
  theme_bw()
```

```{r}
dd_by_row <- jscrape %>%
  mutate(is_dd = as.integer(is_dd),
         # center at 10 words/clue (my wordplay cutoff)
         row = row - 3,
         row_quad = abs(row),
         column = column - 1,
         season = season - 35) %>%
  group_by(web_id, round, column) %>%
  mutate(category_length = if_else(mean(nspaces_clue) > 10,
                                      "long",
                                      "short",
                                      missing = "short")) %>%
  ungroup() %>%
  filter(!is.na(is_dd)) %>%
  nest(-round) %>%
  mutate(model = map(data, ~glm(is_dd ~ factor(row) + category_length + season - 1, 
                                family = binomial(),
                                data = .)))
```

```{r}
summary(dd_by_row$model[[2]])
```

```{r}
summary(dd_by_row$model[[1]])
```

```{r glm of prob of wordplay by column location}
category_length_by_col <- jscrape %>%
  group_by(season, web_id, round, column) %>%
  # signed so 1 is a short category
  summarize(category_length = mean(nspaces_clue, na.rm = T) <= 10) %>%
  ungroup() %>%
  filter(!is.na(category_length)) %>%
  mutate(season = season - 35) %>%
  nest(-round) %>%
  mutate(model = map(data, ~glm(category_length ~ factor(column) + season - 1, family = binomial(), data = .)))

category_length_by_col$model %>% map(summary)
```

Wordplay categories are less likely to be in the 1st column and more likely to be in the 6th column.

```{r}
dd_by_col <- jscrape %>%
  mutate(is_dd = as.integer(is_dd),
         # center at 10 words/clue (my wordplay cutoff)
         season = season - 35) %>%
  filter(!is.na(is_dd)) %>%
  group_by(season, web_id, round, column) %>%
  summarize(category_length = if_else(mean(nspaces_clue, na.rm = TRUE) > 10,
                                      "long",
                                      "short"),
            contains_dd = sum(is_dd)) %>%
  ungroup() %>%
  nest(-round) %>%
  mutate(model = map(data, ~glm(contains_dd ~ factor(column) + category_length + season, 
                                family = binomial(),
                                data = .)))
```

```{r}
summary(dd_by_col$model[[2]])
```

```{r}
summary(dd_by_col$model[[1]])
```

```{r}
dd_by_col %>%
  select(-model) %>%
  unnest() %>%
  filter(contains_dd == 1)

# This is category long-ness by column, categorically
jscrape %>%
  group_by(season, web_id, round, column) %>%
  # signed so 1 is a short category
  summarize(category_length = mean(nspaces_clue, na.rm = T) <= 10) %>%
  ungroup() %>%
  filter(!is.na(category_length)) %>%
  mutate(season = season - 35) %>%
  glm(category_length ~ factor(column) + season - 1, family = binomial(), data = .) %>%
  summary()
```
